---
title: "ML_B_ex1"
author: "itamar_living in dream, Yishay is donkey"
date: "29 3 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

here some intrduction
```{r setup, include=FALSE}
library(tidyverse)
```

import the dataset to a local variable:
```{r}
surveys <- read_csv("netflix-rotten-tomatoes-metacritic-imdb.csv")
```

taking a quick look:
```{r}
str(surveys)
view(surveys)
```

Some information is already known to us just from calling the table. We know how many rows and columns there are, and we have some idea regarding the type of variables (text, numeric) but not much else.

A few commands to help us learn more about the structure of the data are shown here:

```{r}
dim(surveys)
nrow(surveys) 
ncol(surveys) 

head(surveys)
tail(surveys)

names(surveys)

str(surveys)
summary(surveys)
```

## Data Visualisation - It is not working on this dataset!!

```{r}
ggplot(data = surveys) +
  geom_bar(mapping = aes(x = Title))

ggplot(data = surveys) +
  geom_histogram(mapping = aes(x = "Hidden Gem Score"), binwidth = 0.5)

ggplot(data = surveys) +
  geom_histogram(mapping = aes(x = "Hidden Gem Score"), binwidth = 0.1)

ggplot(data = surveys) + 
  geom_bar(mapping = aes(x = Title, fill = Title))

ggplot(data = surveys) + 
  geom_bar(mapping = aes(x = Title, fill = Genre))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
