---
title: "Tirgul 6 - Random Forests"
author: "Romi Goldner Kabeli"
date: "29/04/2022"
output: html_document
---

## Introduction

Today we'll focus on evaluating, tuning and improving our models. This exercise is adapted from https://www.tidymodels.org/start/resampling/ and https://www.tidymodels.org/start/tuning/.

We begin by loading the required libraries:

```{r setup}
library(tidymodels) # for the rsample package, along with the rest of tidymodels
library(modeldata)  # for the cells data
library(vip)        # for variable importance plots
```

We'll be working on the cells dataset. Since this is not exploratory data analysis, we are not going in depth to the data itself, but its always important to at least take a quick at it.

```{r}
# load cells data and look at it
data(cells, package = "modeldata")
cells
```

We have data for 2019 cells, with 58 variables. The main outcome variable of interest for us here is called class, which you can see is a factor. A more detailed explanation can be found at https://www.tidymodels.org/start/resampling/, but for now, lets assume that each row is a cell, classified as "poorly segmented" ("PS") or "well-segmented" ("WS). If we can predict these labels accurately, the larger data set can be improved by filtering out the cells most likely to be poorly segmented.

An important aspect when performing prediction is the class balance, lets take a look:

```{r}
cells %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
```

We have 64% 'poorly segmented' cells and 36% 'well segmented' cells. These groups are not evenly distributed
and we have to deal with by making the splits with the same proportions as the original data.

## Data Splitting

Before we split into test and train sets, we'll remove the "case" variable using the initial_split() command (part of the rsample package), allowing us to take the original data and using its internal split for the partitions.
Note that we are also using the strata argument, which conducts a stratified split. This ensures that, despite the imbalance we noticed in our class variable, our training and test data sets will keep roughly the same proportions of poorly and well-segmented cells as in the original data.

```{r}
set.seed(123)
cell_split <- initial_split(cells %>% select(-case), 
                            strata = class)

cell_train <- training(cell_split)
cell_test  <- testing(cell_split)

nrow(cell_train)
nrow(cell_train)/nrow(cells)

# training set proportions by class
cell_train %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))

# test set proportions by class
cell_test %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))
```

## Random Forest Modeling

The random Forest algorithm uses many decision trees. For classification, which is what
we are focusing on, the final class result is based on the class that was picked from most trees. 

First we define our model with the parsnip package. Random forests have very little mandatory parameters. 
Here we only define the number of trees:

```{r}
# create random forest model
rf_mod <- 
  rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>%   # engine - method of estimation the model will use
  set_mode("classification")
```

Remember to set.seed if you want reproducible results. Taking a look at the model:

```{r}
# train model
set.seed(234)
rf_fit <- 
  rf_mod %>% 
  fit(class ~ ., data = cell_train)
rf_fit
```

We can see our model had 1000 trees like we specified, and several other metrics, such as the prediction error, 12% here.

## Estimating performance

Next we want to see if we can improve our model. We do this by changing some things about it. 
The changes can either be random (guessing) or more precise, depending on performance metrics. In our example, we will use the area under the Receiver Operating Characteristic (ROC) curve (which demonstrates the trade-off between the sensitivity and and specificity), and overall classification accuracy.

Using the yardstick package, let's calculate ROC and Accuracy. Notice that we are still only working with the cell_train partition of our data:

```{r}
rf_training_pred <- 
  predict(rf_fit, cell_train) %>% 
  bind_cols(predict(rf_fit, cell_train, type = "prob")) %>% 
  # Add the true outcome data back in
  bind_cols(cell_train %>% 
              select(class))

# ROC calculation
rf_training_pred %>%                # training set predictions
  roc_auc(truth = class, .pred_PS)

# Accuracy calculation
rf_training_pred %>%                # training set predictions
  accuracy(truth = class, .pred_class)


```

As we can see, these are very good results. Almost too good....Lets see how the model performs on the test data:

```{r}
rf_testing_pred <- 
  predict(rf_fit, cell_test) %>% 
  bind_cols(predict(rf_fit, cell_test, type = "prob")) %>% 
  bind_cols(cell_test %>% select(class))

# ROC calculation
rf_testing_pred %>%                   # test set predictions
  roc_auc(truth = class, .pred_PS)

# Accuracy calculation
rf_testing_pred %>%                   # test set predictions
  accuracy(truth = class, .pred_class)
```

These validation results are lower than the ones we got on the training data.
There are several reasons why training set statistics like the ones shown in this section can be unrealistically optimistic:

- Overfitting - Models like random forests, neural networks, and other black-box methods can essentially memorize the training set. Re-predicting that same set should always result in nearly perfect results.

- The training set does not have the capacity to be a good arbiter of performance. It is not an independent piece of information; predicting the training set can only reflect what the model already knows.

To understand that second point better, think about an analogy of teaching. Suppose you give a class a test, then give them the answers, then provide the same test. The student scores on the second test do not accurately reflect what they know about the subject; these scores would probably be higher than their results on the first test.

## Resampling

Resampling methods, such as cross-validation and the bootstrap, are empirical simulation systems. They create a series of data sets similar to the training/testing split discussed previously; a subset of the data is used for creating the model and a different subset is used to measure performance. Resampling is always used with the training set. 

Here we'll use 10-fold cross-validation. This means that we'll create 10 "mini" datasets, or folds. We call the majority part of the folds (9 out of 10 in this case) the "analysis set" and the minority the "assessment set". We then train a model using the analysis set, and test it on the assessment set, effectively repeating the modeling process 10 times. This is how its done:

```{r}
# create the folds
set.seed(345)
folds <- vfold_cv(cell_train, v = 10)
folds

# here we use a workflow, which bundles together the model specifications and actual modeling
# set the random forest workflow 
rf_wf <- 
  workflow() %>%
  add_model(rf_mod) %>%
  add_formula(class ~ .)

# add folds to workflow and train model
set.seed(456)
rf_fit_rs <- 
  rf_wf %>% 
  fit_resamples(folds)

rf_fit_rs

collect_metrics(rf_fit_rs)
```

We see these results are lower and look more realistic.
Now, looking at the test set for results, we expect and see much more similar results to the above ROC and accuracy:

```{r}
# ROC calculation on test predictions
rf_testing_pred %>%             
  roc_auc(truth = class, .pred_PS)

# Accuracy calculation on test predictions
rf_testing_pred %>%           
  accuracy(truth = class, .pred_class)
```

## Tuning hyperparameters

Another way to improve the performance of our models is by changing the parameters we provide them with. This is also called Tuning. Random Forests, as mentioned above, are not very sensitive to such parameters, but decision trees are. Lets see:

```{r}
# defining the parameters for the decision tree
tune_spec <- 
  decision_tree(
    cost_complexity = tune(), # to control the size of the tree
    tree_depth = tune()
  ) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

tune_spec
```

Note that in the above code, tune() is still just a placeholder. 
We will fill in values later on.

Next, we will create several smaller datasets, to try different parameters. 
The grid_regular command below will create 5 such datasets for each parameter combination, meaning 25 in total.

```{r}
# create 5 'datasets' with different combinations
tree_grid <- grid_regular(cost_complexity(), 
                          tree_depth(),
                          levels = 5)

tree_grid

tree_grid %>% 
  count(tree_depth)

set.seed(234)
cell_folds <- vfold_cv(cell_train) # create the actual cross-validation folds
```

## Model tuning with a grid

```{r}
set.seed(345)
# define the workflow
tree_wf <- workflow() %>%
  add_model(tune_spec) %>%
  add_formula(class ~ .)

# update and train workflow with grids
tree_res <- 
  tree_wf %>% 
  tune_grid(
    resamples = cell_folds,
    grid = tree_grid
    )

tree_res

tree_res %>% 
  collect_metrics()
```

Its easier to see how the models did with a graph:

```{r}
# plot preformance
tree_res %>%
  collect_metrics() %>%
  mutate(tree_depth = factor(tree_depth)) %>%
  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)
```

We can see that our tree with a depth of 1, is the worst model according to both metrics and across all candidate values of cost_complexity. Our deepest tree, with a depth of 15, did better. However, the best tree seems to be between these values with a tree depth of 4. The show_best() function shows us the top 5 candidate models by default:

```{r}
# best tree
tree_res %>%
  show_best("roc_auc")
```

And finally, let's finalize the workflow with the best tree. 

```{r}
best_tree <- tree_res %>%
  select_best("roc_auc") # Selecting the best tree

best_tree

# Finalize workflow
final_wf <- 
  tree_wf %>% 
  finalize_workflow(best_tree)

final_wf
```

## Exploring results

So we have our best tree parameters. Lets take a better look:

```{r}
final_tree <- 
  final_wf %>%
  fit(data = cell_train) 

final_tree
```

Another way to look at how the model uses the different features is using the vip package:

```{r}
final_tree %>% 
  pull_workflow_fit() %>% 
  vip()
```

## Finalizing the model

Finally, let’s return to our test data and estimate the model performance we expect to see with new data. We can use the function last_fit() with our finalized model; this function fits the finalized model on the full training data set and evaluates the finalized model on the testing data.

```{r}
# Set workflow
final_fit <- 
  final_wf %>%
  last_fit(cell_split) 

# train
final_fit %>%
  collect_metrics()

# collect metrics
final_fit %>%
  collect_predictions() %>% 
  roc_curve(class, .pred_PS) %>% 
  autoplot()

args(decision_tree)
```








